I owe thanks to many people for supporting me throughout my
particularly lengthy graduate school endeavor. First and foremost, my
advisor Zhong Shao, who has worked diligently to keep me on the
graduating track. Whenever I became discouraged or disillusioned
with my work, he would always have fresh ideas on how to make it
more exciting. He has a remarkable ability to visualize the ideal destination
of where research should be heading; this was extremely helpful 
whenever I found myself bogged down in gory technical details and Coq proofs.

I am thankful to all of my friends throughout grad school, both
for being colleagues to discuss research with, and for being
companions to relax and play games with. I especially thank 
my friend and officemate Shu-Chun Weng, who always seemed to have 
a clean solution every time I ran into difficulty with Coq proofs.

I am also grateful to my loving family: my parents, siblings, nieces, 
and nephew, who have 
always been supportive of my decision to pursue graduate school~--- despite 
the fact that it requires me being on the opposite side of the country. I 
look forward to finally joining you all in California!

Finally, I thank all of the colleagues in Yale's FLINT group that I have had
the pleasure of working with throughout the years. I am particularly thankful
to Ronghui Gu and Newman Wu, who conducted the bulk of the effort in getting 
an initial verified version of the mCertiKOS kernel up and running. This
initial version was a crucial testbed for demonstrating the applicability
and usability of my new security verification framework; without it, I would
only be able to hope that my theory works well in practice.
